The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-02 manuscript.Rmd
GitLab/Shinyproxy learnings added

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Another example is the package \CRANpkg{golem} ([https://github.com/ThinkR-open/golem](https://github.com/ThinkR-open/golem)), which makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for building and deploying production-grade Shiny applications.
\CRANpkg{googleComputeEngineR} ([https://cloudyr.github.io/googleComputeEngineR/](https://cloudyr.github.io/googleComputeEngineR/)) enables quick deployments of key R services, such as RStudio and Shiny, onto cloud virtual machines (VMs) with Google Cloud Compute Engine \citep{googleComputeEngineR_2019}.
The package utilises `Dockerfile`s to move the labour of setting up those services from the user to a premade Docker image, which is configured and run in the cloud VM.
For example, by specifying the template `template="rstudio"` in functions `gce_vm_template()/gce_vm()` an up-to-date RStudio Server image is launched for development work, whereas specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.

>>>>

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy on virtual machines
can be achieved e.g. via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html). 
A pipeline can include building and checking R packages and Shiny apps. After the code has passed
checks, Docker images are built and pushed to GitLab's container registry
(public or private, depending on project settings). The pipeline is finished off by sending an HTTP
request to the virtual machine where and endpoint (webhook) is waiting to execute the
deployment script (updating configurations, pulling new Docker images, restarting services).

Another example is the package \CRANpkg{golem} ([https://github.com/ThinkR-open/golem](https://github.com/ThinkR-open/golem)), which makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for building and deploying production-grade Shiny applications.
\CRANpkg{googleComputeEngineR} ([https://cloudyr.github.io/googleComputeEngineR/](https://cloudyr.github.io/googleComputeEngineR/)) enables quick deployments of key R services, such as RStudio and Shiny, onto cloud virtual machines (VMs) with Google Cloud Compute Engine \citep{googleComputeEngineR_2019}.
The package utilises `Dockerfile`s to move the labour of setting up those services from the user to a premade Docker image, which is configured and run in the cloud VM.
For example, by specifying the template `template="rstudio"` in functions `gce_vm_template()/gce_vm()` an up-to-date RStudio Server image is launched for development work, whereas specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.
